---
title: "`r params$title`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
params:
  ref_se_path: NULL
  test_se_path: NULL
  title: "Regression Comparison"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 6, fig.height = 5)
library(SummarizedExperiment)
library(ggplot2)
```

```{r load-data}
stopifnot(!is.null(params$ref_se_path), file.exists(params$ref_se_path))
stopifnot(!is.null(params$test_se_path), file.exists(params$test_se_path))

ref_se <- readRDS(params$ref_se_path)
test_se <- readRDS(params$test_se_path)
```

# Overview

Comparing **reference** (`r basename(params$ref_se_path)`) against **test** (`r basename(params$test_se_path)`).

| | Reference | Test |
|---|---|---|
| Rows | `r nrow(ref_se)` | `r nrow(test_se)` |
| Columns | `r ncol(ref_se)` | `r ncol(test_se)` |
| Assays | `r paste(assayNames(ref_se), collapse=", ")` | `r paste(assayNames(test_se), collapse=", ")` |

```{r rowdata-cols}
ref_rd_cols <- colnames(rowData(ref_se))
test_rd_cols <- colnames(rowData(test_se))
rd_shared <- intersect(ref_rd_cols, test_rd_cols)
rd_ref_only <- setdiff(ref_rd_cols, test_rd_cols)
rd_test_only <- setdiff(test_rd_cols, ref_rd_cols)

col_overlap <- data.frame(
  Category = c("Shared", "Reference only", "Test only"),
  Count = c(length(rd_shared), length(rd_ref_only), length(rd_test_only))
)
knitr::kable(col_overlap, caption = "rowData column overlap")
```

```{r rowdata-diffs, results='asis'}
if (length(rd_ref_only) > 0) {
  cat("\n**Reference-only rowData columns:**", paste0("`", rd_ref_only, "`", collapse = ", "), "\n")
}
if (length(rd_test_only) > 0) {
  cat("\n**Test-only rowData columns:**", paste0("`", rd_test_only, "`", collapse = ", "), "\n")
}
```

# Row ID Overlap

```{r venn}
ref_ids <- rownames(ref_se)
test_ids <- rownames(test_se)
shared <- intersect(ref_ids, test_ids)
ref_only <- setdiff(ref_ids, test_ids)
test_only <- setdiff(test_ids, ref_ids)

overlap_df <- data.frame(
  Category = c("Shared", "Reference only", "Test only"),
  Count = c(length(shared), length(ref_only), length(test_only))
)
knitr::kable(overlap_df, caption = "Row ID overlap between reference and test SE objects.")
```

```{r venn-details, results='asis'}
if (length(ref_only) > 0 && length(ref_only) <= 20) {
  cat("\n**Reference-only IDs:**", paste0("`", ref_only, "`", collapse = ", "), "\n")
}
if (length(test_only) > 0 && length(test_only) <= 20) {
  cat("\n**Test-only IDs:**", paste0("`", test_only, "`", collapse = ", "), "\n")
}
```

```{r restrict-shared}
# Restrict to shared rows for all subsequent comparisons
ref_se <- ref_se[shared, ]
test_se <- test_se[shared, ]
```

# Assay: transformedData {.tabset}

Per-sample scatterplots of normalized abundances (reference vs test) for shared proteins.

```{r scatter-fn}
scatter_with_cor <- function(x, y, xlab, ylab, main) {
  both_valid <- !is.na(x) & !is.na(y)
  r <- if (sum(both_valid) > 2) cor(x[both_valid], y[both_valid]) else NA
  df <- data.frame(ref = x, test = y)
  ggplot(df, aes(x = ref, y = test)) +
    geom_point(alpha = 0.3, size = 1) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    labs(x = xlab, y = ylab,
         title = main,
         subtitle = sprintf("r = %.4f  (n = %d)", r, sum(both_valid))) +
    theme_minimal()
}
```

```{r assay-scatter, results='asis', fig.width=5, fig.height=5}
assay_name <- if ("transformedData" %in% assayNames(ref_se)) "transformedData" else assayNames(ref_se)[1]
ref_mat <- assay(ref_se, assay_name)
test_mat <- assay(test_se, assay_name)

shared_samples <- intersect(colnames(ref_mat), colnames(test_mat))

for (samp in shared_samples) {
  cat(sprintf("\n## %s\n\n", samp))
  p <- scatter_with_cor(
    ref_mat[, samp], test_mat[, samp],
    xlab = "Reference", ylab = "Test",
    main = paste0(assay_name, ": ", samp)
  )
  print(p)
  cat("\n\n")
}
```

# Contrast Results {.tabset}

```{r contrast-setup}
ref_rd <- rowData(ref_se)
test_rd <- rowData(test_se)

ref_contrast_cols <- grep("^constrast_", colnames(ref_rd), value = TRUE)
test_contrast_cols <- grep("^constrast_", colnames(test_rd), value = TRUE)
shared_contrasts <- intersect(ref_contrast_cols, test_contrast_cols)

stat_cols <- c("diff", "statistic", "p.value", "FDR")
```

```{r contrast-scatter, results='asis', fig.width=5, fig.height=5}
for (cc in shared_contrasts) {
  cat(sprintf("\n## %s {.tabset}\n\n", cc))

  ref_df <- as.data.frame(ref_rd[[cc]])
  test_df <- as.data.frame(test_rd[[cc]])

  for (sc in stat_cols) {
    if (sc %in% colnames(ref_df) && sc %in% colnames(test_df)) {
      cat(sprintf("\n### %s\n\n", sc))
      p <- scatter_with_cor(
        as.numeric(ref_df[[sc]]), as.numeric(test_df[[sc]]),
        xlab = paste("Reference", sc), ylab = paste("Test", sc),
        main = paste0(cc, ": ", sc)
      )
      print(p)
      cat("\n\n")
    }
  }
}
```

# Summary

```{r summary-table}
# Compute overall correlations for each assay and contrast statistic
summary_rows <- list()

# Assay correlations
for (an in intersect(assayNames(ref_se), assayNames(test_se))) {
  rv <- as.numeric(assay(ref_se, an))
  tv <- as.numeric(assay(test_se, an))
  both <- !is.na(rv) & !is.na(tv)
  r <- if (sum(both) > 2) cor(rv[both], tv[both]) else NA
  summary_rows[[length(summary_rows) + 1]] <- data.frame(
    Source = an, Column = "all samples", r = round(r, 6), n = sum(both)
  )
}

# Contrast correlations
for (cc in shared_contrasts) {
  ref_df <- as.data.frame(ref_rd[[cc]])
  test_df <- as.data.frame(test_rd[[cc]])
  for (sc in stat_cols) {
    if (sc %in% colnames(ref_df) && sc %in% colnames(test_df)) {
      rv <- as.numeric(ref_df[[sc]])
      tv <- as.numeric(test_df[[sc]])
      both <- !is.na(rv) & !is.na(tv)
      r <- if (sum(both) > 2) cor(rv[both], tv[both]) else NA
      summary_rows[[length(summary_rows) + 1]] <- data.frame(
        Source = cc, Column = sc, r = round(r, 6), n = sum(both)
      )
    }
  }
}

summary_df <- do.call(rbind, summary_rows)
knitr::kable(summary_df, caption = "Pearson correlation summary (reference vs test).", row.names = FALSE)
```
